---
title: "FINAL PREDICTION"
author: "Yuang Guo"
date: "2023-11-01"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#ENV PREP
```{r}
  rm(list = ls()) # Clear all files from your environment
  gc()            # Clear unused memory
  cat("\f")       # Clear the console
  graphics.off()  # Clear all graphs
```

```{r}
packages <- c("psych",       # quick summary stats for data exploration,
              "stargazer",   # summary stats,
              "summarytools",# summary stats,
              "naniar",      # for visualisation of missing data,
              "visdat",      # for visualisation of missing data,
              "VIM",         # for visualisation of missing data,
              "DataExplorer",# for visualisation of missing data,
              "tidyverse",   # data manipulation like selecting variables,
              "fastDummies", # Create dummy variables using fastDummies,
              "corrplot",    # correlation plots,
              "ggplot2",     # graphing,
              "data.table",  # reshape for graphing, 
              "car",
              "seasonal",
              "fpp3",
              "tsibble",    # vif for multicollinearity
              "cli",
              "crayon",
              "dplyr",
              "fable",
              "fabletools",
              "feasts",
              "lubridate",
              "magrittr",
              "purrr",
              "rstudioapi",
              "tibble",
              "tidyr",
              "tsibbledata",
              "latex2exp",
              "GGally",
              "broom",
              "zoo",
              "caret",
              "xts",
              "lubridate",
              "stlplus",
              "ISOweek"
              )


for (i in 1:length(packages)) {
  if (!packages[i] %in% rownames(installed.packages())) {
    install.packages(packages[i]
                     , repos = "http://cran.rstudio.com/"
                     , dependencies = TRUE
                     )
  }
  library(packages[i], character.only = TRUE)
}

rm(packages)
```

#import data
```{r}
setwd(getwd())
rawdata = read.csv("petfood_retail_table.csv")
```

#create tsibble for FMCG
```{r}
manu_fmcg <- rawdata |> filter(MANU_ID == "MARS") |> 
                        mutate(DATE = yearmonth(DATE)) |>
                        mutate(transaction = PRICE*UNITS) |>  
                        group_by(DATE) |>
                        summarise(transaction = sum(transaction))

manu_fmcg_monthly_tsibble <- as_tsibble(manu_fmcg, index = "DATE")
manu_fmcg_monthly_tsibble
```

##gg_tsdisplay for original data
```{r}
manu_fmcg_monthly_tsibble |> filter(!is.na(transaction)) |>                                                                                gg_tsdisplay(transaction,plot_type = "partial")
```
##gg_tsdisplay for first_differencing
```{r}
manu_fmcg_monthly_tsibble |> filter(!is.na(transaction)) |> 
                             mutate(transaction = difference(transaction))  |>                                              gg_tsdisplay(transaction,plot_type = "partial")
```
the two cuts off means arima ACF while AT 1,3  means MA 13 PACF at 12 then 1,2


##HERE no need for box-cox transformation because no heteroscadesticity here!!

##decompostion for the original data
```{r}
monthly_dcmp <- manu_fmcg_monthly_tsibble |> filter(!is.na(transaction)) |>
                model(stl = STL(transaction))
components(monthly_dcmp) |> autoplot() + labs(x = "Year-Month", y = "transaction")
```
##check the seasonality
```{r}
Box.test(manu_fmcg_monthly_tsibble$transaction, type = "Ljung-Box")
```
You should add seasonality in the prediction

##autoarima model
```{r}
model <- forecast::auto.arima(manu_fmcg_monthly_tsibble, stepwise = FALSE, approximation = FALSE, seasonal = TRUE)
```

##summary of that model
```{r}
summary(model)
```
AR: yt - yt-1 = φ1(yt-1 - yt-2) + φ2(yt-2 - yt-3) + et
    et =  yt - yt-1 - φ1(yt-1 - yt-2) - φ2(yt-2 - yt-3)

SAR[12]: yt = et - Φ1B12yt     yt = et - Φ1 yt-12    
    et = yt +  Φ1 yt-12
    
Expaination of this model: (1 - φ1B - φ2B2)(1 - B)(1 - Φ1B12)yt = et
φ1：-0.9249     φ2：-0.5523   Φ1：-0.3537
This is our prediction function!
This type CANNOT be transformed to ETS

##summary model for arima 210
```{r}
#model_1 <- ARIMA(manu_fmcg_monthly_tsibble, order = c(2,1,0))
model_1 <- arima(manu_fmcg_monthly_tsibble$transaction , order = c(2,1,0))
model_1
```

        
##Confidence Interval for Forecast
```{r}
forecast_arima <- forecast(forecast::auto.arima(manu_fmcg_monthly_tsibble, stepwise = FALSE, approximation = FALSE, seasonal = TRUE), h = 24)
forecast_arima
```

##Create teh prediction graph
```{r}
manu_fmcg_monthly_tsibble |>
    model(auto_arima = ARIMA(transaction ~ pdq(2,1,0) + PDQ(1,0,0,12))) |>
    forecast(h = 24) |>
    autoplot(manu_fmcg_monthly_tsibble) +
    labs(
         title = "Prediction for FMCG in 2 years from 2023 JAN to 2024 DEC",
         x = "Year-Month",
         y = "total transaction per month ($)",
         )
```


##metrics(using cv)
from that we have 48 period of data, what if we add cv in R,
but I know it's very easy in python
I think LOOCV is better
#I will turn to python for the prediction

##write the data out for python
```{r}
# Export data to a CSV file using write.table
file_path <- "C:\\Users\\guoyy\\OneDrive\\Desktop\\Program\\Mars_petcare\\me\\ts_cv.csv"
write.table(manu_fmcg_monthly_tsibble, 
            file = file_path, 
            sep = ",", col.names = TRUE, row.names = FALSE)

```

##add some comparision based on AIC
```{r}
arima_fit <- manu_fmcg_monthly_tsibble |> 
             filter(!is.na(transaction)) |> 
             model(
                   arima110 = ARIMA(transaction ~ pdq(1,1,1)),
                   arima013 = ARIMA(transaction ~ pdq(1,1,3)),
                   arima011 = ARIMA(transaction ~ pdq(2,1,1)),
                   arima011 = ARIMA(transaction ~ pdq(2,1,3)),
                   SARIMA   = ARIMA(transaction ~ pdq(2,1,0) + PDQ(1,0,0,12)),
                   ETS_best = ETS(transaction ~ error("A") + trend("Ad") + season("N"))
                   )
#here you should be careful that the arima will automatically generate the seasonal part itself
arima_fit
glance(arima_fit) |> arrange(AICc) |> dplyr::select(.model:BIC)
```


##random forest(no way)
```{r}
ets_model <- forecast::ets(manu_fmcg_monthly_tsibble$transaction, restrict = TRUE)
print(ets_model)
forecast_ets <- forecast(ets_model, h = 24)
forecast_ets
```
##creating autoplot of ets prediction
```{r}
fit <- manu_fmcg_monthly_tsibble |>
      model(ETS(transaction ~ error("A") + trend("Ad") + season("N")))
fc <- fit |>
      forecast(h = 24)
fc |>
  autoplot(manu_fmcg_monthly_tsibble)
#together we have the con interval on the above
```

##create weekly data set
```{r}
manu_fmcg_week <- rawdata |> filter(MANU_ID == "MARS") |>
                        mutate(DATE = tsibble::yearweek(DATE)) |>
                        mutate(DATE = format(as.POSIXct(DATE), '%Y-W%U')) |>
                        mutate(transaction = PRICE*UNITS) |>  
                        group_by(DATE) |>
                        summarise(transaction = sum(transaction))
#manu_fmcg_week_tsibble <- as_tsibble(manu_fmcg_week)  |>
 # mutate(DATE = as.Date(paste0(DATE, "-1"), format = "%G-W%V-%u"))
#|>
manu_fmcg_week_tsibble <- manu_fmcg_week |> mutate(DATE = yearweek(DATE)) |>
                          as_tsibble(index = DATE)
  
 #             as_tsibble(index = DATE)

#manu_fmcg_weekly_tsibble <- as_tsibble(manu_fmcg_week, index = "DATE")
#manu_fmcg_weekly_tsibble <- df %>%
 # mutate(date = as.Date(paste0(yearweek, "-1"), format = "%G-W%V-%u")) %>%
  #as_tsibble(index = date, key = "yearweek")
#manu_fmcg_weekly_tsibble
```

##week ggdisplay
```{r}
str(manu_fmcg_week)
str(rawdata)
manu_fmcg_week_tsibble |> autoplot(transaction)
#manu_fmcg_week_tsibble |> gg_tsdisplay(transaction, plot_type = "partial")
manu_fmcg_week_tsibble |> filter(!is.na(transaction)) |>                                                                gg_tsdisplay(manu_fmcg_week_tsibble$transaction,plot_type = "partial")
```


```{r}
#model_week<- forecast::auto.arima(manu_fmcg_week_tsibble, stepwise = FALSE, approximation = FALSE, #seasonal = TRUE)
#summary(model_week)
```

```{r}
# Sample data frame with "yearweek" column
df <- data.frame(
  yearweek = c("2022-W01", "2022-W02", "2022-W03"),
  value = c(10, 12, 15)
)

# Convert "yearweek" to a tsibble
tsib <- df %>%
  mutate(date = as.Date(paste0(yearweek, "-1"), format = "%G-W%V-%u")) %>%
  as_tsibble(index = date, key = "yearweek")

auto_model <- tsib %>%
  model(ARIMA = forecast::auto.arima(value))
```

```{r}
#manu_fmcg_week_fre <- rawdata |> filter(MANU_ID == "MARS") |>
 #                       mutate(DATE = format(as.POSIXct(DATE), '%Y-W%U')) |>
    #                   mutate(DATE = yearmonth(DATE)) |>  
     #                 mutate(DATE = isoweek(DATE)) |>
      ###                summarise(transaction = sum(transaction))
#myts <- ts(manu_fmcg_week_fre, start = c(""), frequency = 52)
#myts
```

```{r}
#current_date <- Sys.Date()
#current_year_week <- lubridate::isoweek(current_date)
#print(current_year_week)
```


#tips, exponential smoothing(ETS is different from ARIMA in this model) and random forest
we need to compare both AIC, AICc and BIC here.  TSCV already put it in python
trying to explain the overlap of that ETS and ARIMA the map from fpp3
#economics <- ggplot2::economics
$knitr::kable(head(economics)